### 2016-11-18 bio.tex
Changed names and all for FGCS

Pablo García-Sánchez received his Degree (2007), MSc degree (2008) and PhD (2014) in computer science and engineering at the University of Granada (Spain). He is currently a substitute professor in the University of Cádiz. He has participated and lead several national research projects related to e-Health, web-services, videogames and optimization, whose results have been published in conferences (such as GECCO, PPSN or EVO*) and international journals (such as Applied Soft Computing). Also, he has been a visitor researcher to the Vrije Universiteit Amsterdam. His interests include service-oriented computing, evolutionary computation, computational intelligence in videogames and distributed algorithms.