### 2016-11-18 highlights.tex

\documentclass{article}
\usepackage{graphicx}

\begin{document}

%\title{Introduction to \LaTeX{}}
%\author{Author's Name}

%\maketitle

%\section{Reviewer 1}

\begin{itemize}
\item A study on methods to overlap the search space in distributed multi-objective evolutionary algorithms is carried out
\item A new adaptive overlapping mechanism is presented
\item Different combinations of number of islands, genome sizes, benchmarks and methods are used
\item Different quality metrics have been compared
\item Results are validated in a real system
\item All overlapping schemes outperform the baseline, being the adaptive method the one that obtains the best performance
\end{itemize}
\end{document}