### 2019-05-07 highlights.tex
Changes in the highlights

\documentclass{article}
\usepackage{graphicx}

\begin{document}

%\title{Introduction to \LaTeX{}}
%\author{Author's Name}

%\maketitle

%\section{Reviewer 1}

\begin{itemize}
\item ------------------------------------------------------------------------------------
\item A study on methods to overlap the search space in distributed MOEAs is carried out  
\item A new automatic overlapping mechanism is presented
\item Results are validated in a real system, using different quality metrics
\item Relation between the number of islands and chromosome length is a relevant factor
\item Best results are obtained when increasing the dimensionality and overlapping size
%\item A study on methods to overlap the search space in distributed multi-objective evolutionary algorithms is carried out
%\item A new adaptive overlapping mechanism is presented
%\item Different combinations of number of islands, chromosome sizes, benchmarks and methods are used
%\item Different quality metrics have been compared
%\item Results are validated in a real system
%\item Results show that the relation between the number of islands (subpopulations) and the length of the chromosome (number of decision variables) is a relevant factor to determine the most efficient alternative to distribute the decision variables
%\item The higher the dimensionality and the overlapping, the greater the benefit of distribution in a large number of islands
\end{itemize}
\end{document}